---
title: "Project-1, ST558"
author: "Pratap Adhikari"
date: "9/18/2020"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '1'
    toc_float: true
---

```{r setup, include=FALSE, cache=T, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
library(haven)
library(ggplot2)
#install.packages("qwraps2")
library(qwraps2)
library(rmarkdown)
library(RSQLite)
```

### Function to get franchiseAPI
```{r}

nhl<- function(tabName, ID=NULL, ...){
  base_url<- "https://records.nhl.com/site/api"
  if (!is.null(tabName)){
 
     if ( tabName %in% c("franchise", "franchise-team-totals") && (!is.null(ID))){
    stop("This tab can not return with 'ID' defined")
  }
  
    if (is.null(ID)){ 
    full_url<- paste0(base_url, "/", tabName) 
  }
  
    if (!is.null(ID)){
     full_url<- paste0(base_url, "/", tabName, ID) 
  }
  get_nfl<- GET(full_url)
  txt_nfl<- content(get_nfl, "text")  # convert to JSON text form
  json_nfl<- fromJSON(txt_nfl, flatten=T) # convert to list
return(json_nfl)
  }
  else {
    return("Invalid tabName")
  }
}
```

```{r, echo=F}

#Test of the function
#nhl("franchise")
#nhl("franchise-team-totals")
#nhl("franchise-season-records?cayenneExp=franchiseId=", 15)
#nhl("franchise-goalie-records?cayenneExp=franchiseId=", 15)
#nhl("franchise-skater-records?cayenneExp=franchiseId=", 15)
#nhl("franchise-goalie-records")
``` 

### Function to get statsAPI
```{r}

nhl_modifier<- function(modifier, ID=NULL,...){
  stbase_url<- "https://statsapi.web.nhl.com/api/v1/teams"
  if (modifier %in%  c("expand=team.roster", "expand=person.names", "expand=team.schedule.next", "expand=team.schedule.previous",  "expand=team.stats",  "expand=team.roster&season=20142015", "teamId=4,5,29",  "stats=statsSingleSeasonPlayoffs"))
  {
    get_st<- GET(paste0(stbase_url, "?", modifier))
st_txt<- content(get_st, "text")
json_st<- fromJSON(st_txt, flatten=T)

}
else {
  json_st="Sorry, can't accept this modifier"
  
}
return(json_st)
}
```

```{r, echo=F}
#Test of the function
#nhl_modifier("expand=team.roster")
#nhl_modifier("expand=person.names")
#nhl_modifier("expand=team.schedule.next")
#nhl_modifier("expand=team.schedule.previous")
#nhl_modifier("expand=team.stats")
#nhl_modifier("expand=team.roster&season=20142015")
#hl_modifier("teamId=4,5,29")
#nhl_modifier("stats=statsSingleSeasonPlayoffs")
#nhl_modifier("dstger")
```

```{r}
nhlData<- function (tabName=NULL, modifier=NULL, ID=NULL, ...){
if (!is.null(tabName) && !is.null(modifier)){
  stop("it can not work together with tabName and modifier")
 
}  
if(is.null(tabName) && is.null(modifier) ){
    
  output<- nhl("franchise")
}
  if (is.null(modifier) && !is.null(tabName)){
     output<- nhl(tabName, ID)
     output<- output$data
  }
  # if modifier is not null and id is null
  if(!is.null(modifier) && is.null(id)){
   output<- nhl_modifier(modifier)
  }
  
  #if both modifier and id are not  null
  if(!is.null(modifier) && !is.null(id) ){
    output<- nhl_modifier(modifier)
    output<- output$teams
    output<- output %>% filter(id==ID) %>% select(id:roster.link)
    
  }
  return(output)
}
```

```{r, echo=F}
# Test of the function
#nhlData(modifier = "expand=team.roster", ID=8)
#nhlData(modifier = "expand=team.roster", ID=2
#nhlData(tabName  = "franchise-goalie-records?cayenneExp=franchiseId=", ID=15)
#nhlData(tabName = "franchise")
#nhlData(tabName =  "franchise-team-totals", ID=15)
#nhlData("franchise-team-totals", )
```


### Overview of franchise and location 
```{r}
teamtotal<- nhl(tabName = "franchise-team-totals") 

#getteams from another endpoint

division<- nhl_modifier(modifier = "expand=team.roster")$teams %>% select(id, division.name,  locationName, division.nameShort, conference.name) %>% rename(teamId=id)

#join the two dataset from two different APIs

newData<- left_join(teamtotal, division, by="teamId")
head(newData, n=4)


# overview of after joining two datasets from two different API endpoints
kable(head(newData %>% select(id, franchiseId, teamName, locationName)) , caption= "Franchise ID, Team Name, Location  table for your reference:")

```


Read table from two different APIs
```{r}
dta1<- nhlData("franchise")
dta1<- dta1 %>% select(id, mostRecentTeamId, teamCommonName, teamPlaceName) 

golietable15<- nhl(tabName  = "franchise-goalie-records?cayenneExp=franchiseId=", ID=15)$data %>% select(activePlayer, firstName, franchiseName, gamesPlayed, lastName, gameTypeId, playerId, ties, wins, losses )

```


## Analysis on team ID=15 (Dallas Stars)
```{r}
#Create new variable by adding first and last name from two different columns

golietable15$playerName<- c(paste0(golietable15$firstName, " ", golietable15$lastName))

#select only the varaibles required to analyse the data
golietable15<- golietable15 %>% select(franchiseName, playerName, playerId, activePlayer, gameTypeId, gamesPlayed,ties, wins, losses)
golietable15
```

### Active Players 

Table showing active players from the **Dallas** team

```{r}
 table (golietable15  %>% group_by(activePlayer)%>% select(franchiseName, activePlayer))

```


### Win/Loss Rate

```{r}
#create new variables 
wlRate<- golietable15 %>% mutate(tiesRate= round(ties/gamesPlayed, 2), winRate= round(wins/gamesPlayed, 2) , lossRate= round(losses/gamesPlayed,2), MissedClcn= 1-(tiesRate+winRate+lossRate) ) %>% select(playerName, playerId, gameTypeId, gamesPlayed, activePlayer, winRate, lossRate, MissedClcn)
kable(wlRate)

```


Numeric Summary


### Sumamry table

```{r}
sumry<- function (x, ...){
  dta<- wlRate %>% filter(gameTypeId == x) %>% select(winRate, lossRate)
  if (x==2) type<- "regular season" else type<- "play off season"
  kable (apply(dta, 2, summary), format="html", digit =4, caption = paste0("Summary among all of the players"))
}

# Regular season summary
sumry(2)

```


## Plots
### Scatter 

The higest win rate is the individuals who have played less games. But importently, higher number of games player looks like they are more consistent in winnig rate than the lower number of games player.
```{r}

#scalter plot

plot(golietable15$gamesPlayed, golietable15$wins/golietable15$gamesPlayed , col="blue",
     xlab="Games Played ",
     ylab = "Games won",
     main = "Games played vs win rate by individual players")
```

### Box plot


```{r}
#box plot
bxPlot1<- ggplot(data= golietable15, aes(x=gamesPlayed, y= losses, group=activePlayer, color=activePlayer))
bxPlot1 + geom_boxplot() + labs(title="Boxplot of games played and losses by individual players") + geom_jitter(aes(color=activePlayer))

```

### Bar plots about activeplayers


```{r}
type2<-  golietable15 %>% select(playerName, gamesPlayed, wins, activePlayer, gameTypeId)
barPlot1<- ggplot(data=type2, aes(x=activePlayer))
barPlot1 + geom_bar(aes(fill= activePlayer), position = "dodge") + labs(title = "Bar plot about active/inactive players")

```

### Histogram of games Played 

The histogram for games played grahp shows that there are very few players playing the regular season who are tagged as active player. They seem to be played between 40 to 150 games throught their entire career for the data collected period. Also, it shwos there are large group of players who played less than 5 games.

```{r}
type3<- golietable15 %>% select(wins, losses, activePlayer, gamesPlayed) #(gameTypeId==3,
histogram1<- ggplot(data=type3, aes(x=gamesPlayed))
histogram1 + geom_histogram(binwidth = 1, aes(fill= activePlayer)) + labs(title="Histogram for games Played") +
  geom_density(adjust= 0.25, alpha=0.05)

```

Here we can see the active players to have the highest number of wins.

```{r}
type3<- golietable15 %>% select(wins, losses, activePlayer, gamesPlayed, playerId) #(gameTypeId==3, division.name=="Atlantic") 
histogram1<- ggplot(data=type3, aes(x=wins))
histogram1 + geom_histogram(binwidth = 1, aes(y=..density.., fill= activePlayer )) + labs(title="Histogram for wins") +  geom_density(adjust= 0.2, alpha=0.05)

```

For the shake of try, the above histograms does not look preety well distributed data, so I tried below to see it on win rates, a calculated variable. The win rate is left skewed.

```{r}

hist(wlRate$winRate, probability = T, col = "light blue", xlab = "Win Rate", main = "Histogram of Win Rate (wins/gameplayed)")
#lines(seq(from=0, to=0.7, by = 0.005), dnorm(seq(from=0, to=0.7, by=0.005)), col="Red")
```

```{r}
render("Proj1.Rmd", output_format= "github_document", output_file= I("README.md") )

```










