---
title: "Project-1, ST558"
author: "Pratap Adhikari"
date: "9/18/2020"
output:
  html_document:
    toc: yes
    toc_depth: 5
  pdf_document:
    toc: yes
    toc_depth: '5'
  rmarkdown::github_document:
    toc: yes
always_allow_html: yes
---

```{r setup, echo=F, warning=F, message=F}
library(knitr)
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
library(haven)
library(ggplot2)
library(qwraps2)
library(rmarkdown)
library(RSQLite)
```

## Project1-ST558

This project work involves creating a `vignette` for reading and summarizing data from the *National Hockey League's* (NHL) **API**.


### List of library packages  
The list of library packages I have used to run this code in order to carry on this project are:

* knitr  
* httr  
* jsonlite  
* tidyverse  
* dplyr  
* haven  
* ggplot2  
* qwraps2  
* rmarkdown  
* RSQlite

#### To install the packages:
install.packages("knitr", "httr", "jsonlite", "tidyverse", "dplyr", "haven", "ggplot2", "qwraps2", "rmarkdown", "RSQlite")

### Function to get franchiseAPI


```{r}
#create funciton to read data from records API
nhl<- function(tabName, ID=NULL, ...){
  base_url<- "https://records.nhl.com/site/api"
  if (!is.null(tabName)){
 
     if ( tabName %in% c("franchise", "franchise-team-totals") && (!is.null(ID))){
    stop("This tab can not return with 'ID' defined")
  }
  
    if (is.null(ID)){ 
    full_url<- paste0(base_url, "/", tabName) 
  }
  
    if (!is.null(ID)){
     full_url<- paste0(base_url, "/", tabName, ID) 
  }
  get_nfl<- GET(full_url)
  txt_nfl<- content(get_nfl, "text")  # convert to JSON text form
  json_nfl<- fromJSON(txt_nfl, flatten=T) # convert to list
return(json_nfl)
  }
  else {
    return("Invalid tabName")
  }
}
```

```{r, echo=F}
#Test of the function
#nhl("franchise")
#nhl("franchise-team-totals")
#nhl("franchise-season-records?cayenneExp=franchiseId=", 15)
#nhl("franchise-goalie-records?cayenneExp=franchiseId=", 15)
#nhl("franchise-skater-records?cayenneExp=franchiseId=", 15)
#nhl("franchise-goalie-records")
``` 

### Function to get statsAPI
```{r}
#create funciton to data from statsAPI
nhl_modifier<- function(modifier, ID=NULL,...){
  stbase_url<- "https://statsapi.web.nhl.com/api/v1/teams"
  if (modifier %in%  c("expand=team.roster", "expand=person.names", "expand=team.schedule.next", "expand=team.schedule.previous",  "expand=team.stats",  "expand=team.roster&season=20142015", "teamId=4,5,29",  "stats=statsSingleSeasonPlayoffs"))
  {
    get_st<- GET(paste0(stbase_url, "?", modifier))
st_txt<- content(get_st, "text")
json_st<- fromJSON(st_txt, flatten=T)
}
else { #return a message if the modifier is not campatible with the function
  json_st="Sorry, can't accept this modifier"
  
}
return(json_st)
}
```

```{r, echo=F}
#Test of the function
#nhl_modifier("expand=team.roster")
#nhl_modifier("expand=person.names")
#nhl_modifier("expand=team.schedule.next")
#nhl_modifier("expand=team.schedule.previous")
#nhl_modifier("expand=team.stats")
#nhl_modifier("expand=team.roster&season=20142015")
#hl_modifier("teamId=4,5,29")
#nhl_modifier("stats=statsSingleSeasonPlayoffs")
#nhl_modifier("dstger")
```

```{r}
nhlData<- function (tabName=NULL, modifier=NULL, ID=NULL, ...){
if (!is.null(tabName) && !is.null(modifier)){
  stop("it can not work together with tabName and modifier")
 
}  
if(is.null(tabName) && is.null(modifier) ){
    
  output<- nhl("franchise")
}
  if (is.null(modifier) && !is.null(tabName)){
     output<- nhl(tabName, ID)
     output<- output$data
  }
  # if modifier is not null and id is null
  if(!is.null(modifier) && is.null(id)){
   output<- nhl_modifier(modifier)
  }
  
  #if both modifier and id are not  null
  if(!is.null(modifier) && !is.null(id) ){
    output<- nhl_modifier(modifier)
    output<- output$teams
    output<- output %>% filter(id==ID) %>% select(id:roster.link)
    
  }
  return(output)
}
```

```{r, echo=F}
# Test of the function
#nhlData(modifier = "expand=team.roster", ID=8)
#nhlData(modifier = "expand=team.roster", ID=2
#nhlData(tabName  = "franchise-goalie-records?cayenneExp=franchiseId=", ID=15)
#nhlData(tabName = "franchise")
#nhlData(tabName =  "franchise-team-totals", ID=15)
#nhlData("franchise-team-totals", )
```


### Overview of franchise and location 
```{r}
teamtotal<- nhl(tabName = "franchise-team-totals")$data
#getteams from another endpoint
division<- nhl_modifier(modifier = "expand=team.roster")$teams %>% select(id, division.name,  locationName, division.nameShort, conference.name) %>% rename(teamId=id)
#join the two dataset from two different APIs
newData<- left_join(teamtotal, division, by="teamId")
head(newData, n=4)
```

```{r}
# overview of after joining two datasets from two different API endpoints
kable(newData %>% select(id, franchiseId, teamName, locationName) , caption= "Franchise ID, Team Name, Location  table for your reference:")
```


Read table from two different APIs
```{r}
dta1<- nhlData("franchise")
dta1<- dta1 %>% select(id, mostRecentTeamId, teamCommonName, teamPlaceName) 
golietable15<- nhl(tabName  = "franchise-goalie-records?cayenneExp=franchiseId=", ID=15)$data %>% select(activePlayer, firstName, franchiseName, gamesPlayed, lastName, gameTypeId, playerId, mostGoalsAgainstOneGame, ties, wins, losses )
```


### Analysis on team ID=15 (Dallas Stars)
```{r}
#Create new variable by adding first and last name from two different columns
golietable15$playerName<- c(paste0(golietable15$firstName, " ", golietable15$lastName))
#select only the varaibles required to analyse the data
golietable15<- golietable15 %>% select(franchiseName, playerName, playerId, activePlayer, gameTypeId, gamesPlayed, mostGoalsAgainstOneGame, ties, wins, losses)
```

#### Categorical Summary 

Table showing active players from the **Dallas** team

```{r}
library(kableExtra)
 freqtbl<- table (golietable15  %>% group_by(activePlayer)%>% select(franchiseName, activePlayer))
freqtbl
add_header_above(header = c("Franchise" = 1, "Active Player" = 2),   kable(freqtbl))
```


```{r}
# Max goals against one game
kable( table(golietable15 %>% group_by(mostGoalsAgainstOneGame) %>% select(mostGoalsAgainstOneGame, playerName)))
```


#### Win/Loss Rate

```{r}
#create new variables 
wlRate<- golietable15 %>% mutate(tiesRate= round(ties/gamesPlayed, 2), winRate= round(wins/gamesPlayed, 2) , lossRate= round(losses/gamesPlayed,2), MissedClcn= 1-(tiesRate+winRate+lossRate) ) %>% select(playerName, playerId, gameTypeId, gamesPlayed, activePlayer, winRate, lossRate, MissedClcn)
kable(wlRate)
```





#### Sumamry table
Numeric Summary
```{r}
sumry<- function (x, ...){
  dta<- wlRate %>% filter(gameTypeId == x) %>% select(winRate, lossRate)
  if (x==2) type<- "regular season" else type<- "play off season"
  kable (apply(dta, 2, summary), format="html", digit =4, caption = paste0("Summary among all of the players"))
}
# Regular season summary
sumry(2)
```

Sumamry of number of games played
```{r}
gpl<- golietable15 %>% select(gamesPlayed, wins)
kable (apply(gpl, 2, summary) )
```

#### Plots
Every plot I am trying to show the plot difference and advanced options available to plot the same plots.
##### Scatter 

The higest win rate is the individuals who have played less games. But importently, higher number of games player looks like they are more consistent in winnig rate than the lower number of games player.
```{r}
#scalter plot
plot(golietable15$gamesPlayed, golietable15$wins/golietable15$gamesPlayed , col="blue",
     xlab="Games Played ",
     ylab = "Games won",
     main = "Games played vs win rate")
```


The active players has larger width, which can be the effect of very lower number of players .
```{r}
ggplot (golietable15, aes(x=gamesPlayed, y=wins, group=activePlayer)) + geom_point(aes(color= activePlayer)) + geom_smooth(method='lm', color='light blue') + ggtitle("Games played vs Wins")
```


##### Box plot

The players not tagged as active players looks to have loose the higest number of games. But in the median number of games lost is higher for active players.
```{r}
#box plot
bxPlot1<- ggplot(data= golietable15, aes(x=gamesPlayed, y= losses, group=activePlayer, color=activePlayer))
bxPlot1 + geom_boxplot() + labs(title="Boxplot of games played and losses by individual players") + geom_jitter(aes(color=activePlayer))
```

##### Bar plots 

Again, there are very few, less than 5 players are categorized as active player, which seems unreal.

```{r}
type2<-  golietable15 %>% select(playerName, gamesPlayed, wins, activePlayer, gameTypeId)
barPlot1<- ggplot(data=type2, aes(x=activePlayer))
barPlot1 + geom_bar(aes(fill= activePlayer), position = "dodge") + labs(title = "Bar plot about active/inactive players")
```

##### Histogram

The histogram for games played grahp shows that there are very few players playing the regular season who are tagged as active player. They seem to be played between 40 to 150 games throught their entire career for the data collected period. Also, it shwos there are large group of players who played less than 5 games.the different color is contribution of two categorized playes as active or not.The distribution is extremely right skwed

```{r}
type3<- golietable15 %>% select(wins, losses, activePlayer, gamesPlayed) #(gameTypeId==3,
histogram1<- ggplot(data=type3, aes(x=gamesPlayed))
histogram1 + geom_histogram(binwidth = 10, aes(fill= activePlayer)) + labs(title="Histogram for games Played") +
  geom_density(adjust= 0.25, alpha=0.05)
```
The histogram looks extremly right skewed that the players playing higher number of games are lesser
```{r}
ggplot(golietable15, aes(x=gamesPlayed, ..density..)) + geom_histogram(bins=20) + ggtitle("Histogram for Games Played") + ylab("Density") + geom_density(col="red", lwd=2, adjust=2)
```
Again, the different color is contribution of two categorized playes as active or not.

```{r}
type3<- golietable15 %>% select(wins, losses, activePlayer, gamesPlayed, playerId) #(gameTypeId==3, division.name=="Atlantic") 
histogram1<- ggplot(data=type3, aes(x=wins))
histogram1 + geom_histogram(binwidth = 5, aes(y=..density.., fill= activePlayer )) + labs(title="Histogram for wins") +  geom_density(adjust= 0.3, alpha=0.05)
```

For the shake of try, the above histograms does not look preety well distributed data, so I tried below to see it on win rates, a calculated variable. The win rate is left skewed.

```{r}
hist(wlRate$winRate, probability = T, col = "light blue", xlab = "Win Rate", main = "Histogram of Win Rate (wins/gameplayed)")
```


```{r}
ggplot (wlRate, aes(x=winRate, ..density..)) + geom_histogram(bins=20) + facet_wrap(golietable15$activePlayer) + ggtitle("Histogram for Wins Rate with active player(True/False)") + ylab("Density") + geom_density(col="red", lwd=2, adjust=1)
```
